Transcriptomic Analysis for Prognostic Value in Head and Neck Squamous Cell Carcinoma

### 2021-07-06 Cover_letter_JPM.tex
Updates from Overleaf

@@

% mark in blue or red
\usepackage{xcolor}
\definecolor{asparagus}{rgb}{0.53, 0.66, 0.42}
%\usepackage[table,xcdraw]{xcolor}
\newenvironment{MyColorPar}[1]{%
    \leavevmode\color{#1}\ignorespaces%
}{%
}%

@@

We also think that the sliding-window selection of cutoffs would capture more genes whose \textit{P} values just reach the significant level (\textit{P}=0.05).
%http://bioinformatica.mty.itesm.mx:8080/Biomatec/SurvivaX.jsp run TCGA2016 in median (median cutoff; not case number 50\% 50\%)
%=> defense: 
%% *** 準備一張 20 candidates: accumulated \textit{P}-value plots 來(like Figure 4: (Please see supplementary Figure SXXX))
%說明，其實這些 P values 不論 50% cut or 70% 幾乎都 far less than 0.05
% file: 1453 Feb 15 2020 ./intermediate_data/HNSCC_OS_marginSFP_pvalueBonferroni_KM_candidate_cox.Rda => candidate_sample
%[1] "HNSCC_OS_marginS_pvalueBonferroni_sorted"   
%[2] "HNSCC_OS_marginFree_pvalueBonferroni_sorted"
%[3] "HNSCC_OS_marginPlus_pvalueBonferroni_sorted"
Since we reviewed the results of sliding-window cutoff finding (from 50\% to 70\% quantile), the 20 preliminary candidates have FDR-corrected \textit{P} values from 1e-4 to 0.04 (or 0.001) (*** see a table?).
The top 3 the 3 biomarker candidates, CAMK2N1, CALML5, and FCGBP, have FDR-corrected \textit{P} values from \num{1.63e-5} to \num{1.97e-4} (please see Figure \ref{fig:figure4}, 
%which have been grouped by cutoffs between; 
it is also shown in manuscript as updated "Figure 5").
Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in 
supplementary figure S1). After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233}.
%Rplot_pvaluePlot_NDFIP1.pdf
%  to find cases_OS outside of range 180-220 (== cutoff not at median)
%# pvalue plot as 倒V字形 "^" or W p_OS pvalues plots (with local minials)
% # 99 genes have one p_OS >= 0.05, in OS_pvalueTex005
%# 3641 genes have one p_OS >= 0.02, in OS_pvalueTex002
%# special case: HNSCC_survivalAnalysis_marginS_NDFIP1.xlsx
%# NDFIP1 (within our 20 preliminary candidates), 
In those \textit{P}-value plots, we could see that the sliding-window selection of cutoff finding might collect more potential candidates (\textit{P} values far less than $0.001$) for further Cox's modeling.
%Some useful biomarkers might be kick-out prematurely in "median cutoff" of survival analysis.
Indeed, it needs to determine its effect size (hazard ratio) then to be validated by the other independent cohort.

>>>>

\usepackage{array}
\usepackage{ragged2e}
\usepackage{rotating}
\usepackage{tabularx} % for resizebox?
\usepackage{makecell}
% \usepackage{array}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{siunitx} % for  1e-10 scientific notation
%\usepackage{caption}
%\usepackage{subcaption}
\usepackage{booktabs, multirow} % for borders and merged ranges
\usepackage{soul}% for underlines
%\usepackage[table]{xcolor} % for cell colors
\usepackage{changepage,threeparttable}

@@

% mark in blue or red
\usepackage{xcolor}

\definecolor{asparagus}{rgb}{0.53, 0.66, 0.42}
%\usepackage[table,xcdraw]{xcolor}
\newenvironment{MyColorPar}[1]{%
    \leavevmode\color{#1}\ignorespaces%
}{%
}%

@@

Since the sliding-window selection of cutoffs would capture more genes whose \textit{P} values just reach the significant level (\textit{P} = 0.05),
%http://bioinformatica.mty.itesm.mx:8080/Biomatec/SurvivaX.jsp run TCGA2016 in median (median cutoff; not case number 50\% 50\%)
%=> defense: 
%% *** 準備一張 20 candidates: accumulated \textit{P}-value plots 來(like Figure 4: (Please see supplementary Figure SXXX))
%說明，其實這些 P values 不論 50% cut or 70% 幾乎都 far less than 0.05
% file: 1453 Feb 15 2020 ./intermediate_data/HNSCC_OS_marginSFP_pvalueBonferroni_KM_candidate_cox.Rda => candidate_sample
%[1] "HNSCC_OS_marginS_pvalueBonferroni_sorted"   
%[2] "HNSCC_OS_marginFree_pvalueBonferroni_sorted"
%[3] "HNSCC_OS_marginPlus_pvalueBonferroni_sorted"
we reviewed the results of cutoffs from 50\% to 70\% quantile to find the evidence.
%> View(hazards[hazards$Gene.Symbol %in% c(gene_labeled_silver, gene_labeled_golden), c("Gene.Symbol","FDR.Pvalue")])
A total of 20 preliminary candidates have FDR-corrected \textit{P} values from \num{3.78e-6} to \num{3.03e-4} (please see Table \ref{table:table1} and \ref{table:table3}; also shown in supplementary Table S1, S2).
The top 3 biomarker candidates, CAMK2N1, CALML5, and FCGBP, have FDR-corrected \textit{P} values from \num{1.63e-5} to \num{1.97e-4} (please see Figure \ref{fig:figure4}, 
%which have been grouped by cutoffs between; 
it is also shown in manuscript as updated "Figure 5").
Moreover, the gene NDFIP1, one of the 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff.
NDFIP1 also has a \textit{P} value of \num{2.62e-6} when cut at 70\% quantile  (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in 
supplementary figure S1). 
After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233}.
%Rplot_pvaluePlot_NDFIP1.pdf
%  to find cases_OS outside of range 180-220 (== cutoff not at median)
%# pvalue plot as 倒V字形 "^" or W p_OS pvalues plots (with local minials)
% # 99 genes have one p_OS >= 0.05, in OS_pvalueTex005
%# 3641 genes have one p_OS >= 0.02, in OS_pvalueTex002
%# special case: HNSCC_survivalAnalysis_marginS_NDFIP1.xlsx
%# NDFIP1 (within our 20 preliminary candidates), 
In those \textit{P}-value plots, we could see that the sliding-window selection of cutoff finding might collect more candidates (\textit{P} values far less than $0.001$) for further Cox's modeling.
%Some useful biomarkers might be kick-out prematurely in "median cutoff" of survival analysis.
Indeed, it needs to determine its effect size (hazard ratio) then to be validated by the other independent cohort.

%%% supplementary tables
%Supplementary Table \ref{table:table1}: \\
%Overexpression of 10 genes with Cox's hazard ratio $> 1.5$.\\

%Supplementary Table \ref{table:table3}: \\
%Overexpression of 10 genes with Cox's hazard ratio $< 0.6$. \\

% tables below %%%%
% table1
\begin{table}[ht]
%\begin{sidewaystable}[hp]
%\centering
\raggedright
\caption{ The top 10 genes overexpressed with poor prognosis in \acrshort{hnscc} (ranked by adjusted \textit{P} value) }
\arrayrulecolor[rgb]{0.255,0.255,0.255}
\resizebox{0.6\linewidth}{!}{%
%\begin{tabular}{|l|l|l|l|l|l|l|l|c|}
\begin{tabularx}{\textwidth}{|l|p{6.0cm}|l|l|l|l|l|l|} % *** capital X or p{2cm}
% p{2cm}
\cline{1-8}
\multicolumn{1}{|c|}{\multirow{2}{*}{Gene ID}} & \multicolumn{1}{c|}{\multirow{2}{*}{Gene Description}}     & \multicolumn{2}{c|}{Kaplan-Meier survival}                                                                                   & \multicolumn{2}{c|}{Univariate~}                                                        & \multicolumn{2}{c|}{Multivariate} \\
\cline{3-8}
\multicolumn{1}{|c|}{}                         & \multicolumn{1}{c|}{}                                      & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}FDR \\ \textit{P} value\end{tabular}}               & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Bonferroni \\\textit{P} value\end{tabular}} & \multicolumn{1}{c|}{HR*}                   & 95\% CI                                    & HR*                                        & 95\% CI   \\\cline{1-8}
DKK1                                           & dickkopf WNT signaling pathway inhibitor 1                 & \num{3.8e-6}                               & 0.001                                                                           & 2.266                                      & 1.666-3.082                                & 2.135                                      & 1.559-2.924     \\                          % &                                           \\ 
\cline{1-8}
CAMK2N1                                        & calcium/calmodulin-dependent protein kinase II inhibitor 1 & \num{1.5e-5}                               & 0.002                                                                           & 2.101                                      & 1.572-2.809                                & 2.007                                      & 1.490-2.704  \\                            % &                                           \\ %\checkmark
\cline{1-8}
STC2                                           & stanniocalcin 2                                            & \num{1.5e-5}                               & 0.004                                                                           & 2.147                                      & 1.578-2.921                                & 2.075                                      & 1.515-2.843                             \\%   &                                           \\ 
\cline{1-8}
PGK1                                           & phosphoglycerate kinase 1                                  & \num{2.4e-5}                               & 0.006                                                                           & 2.127                                      & 1.563-2.895                                & 2.046                                      & 1.498-2.795                              \\%    &                                           \\ %\checkmark
\cline{1-8}
SURF4                                          & surfeit 4                                                  & \num{6.2e-5}                               & 0.006                                                                           & 2.055                                      & 1.531-2.757                                & 2.089                                      & 1.543-2.829                               \\%   &                                          \\ %\checkmark  
\cline{1-8}
USP10                                          & ubiquitin specific peptidase 10                            & \num{7.9e-5}                               & 0.012                                                                           & 2.083                                      & 1.532-2.834                                & 2.119                                      & 1.551-2.895                              \\%    &                                          \\ %\checkmark
\cline{1-8}
NDFIP1                                         & Nedd4 family interacting protein 1                         & \num{1.1e-4}                               & 0.017                                                                           & 2.031                                      & 1.502-2.746                                & 2.027                                      & 1.483-2.771                               \\%   &                                           \\ %\checkmark 
\cline{1-8}
FOXA2                                          & forkhead box A2                                            & \num{1.6e-4}                               & 0.018                                                                           & 1.976                                      & 1.479-2.640                                & 1.914                                      & 1.426-2.569                               \\%   &                                         \\ %\checkmark  
\cline{1-8}
STIP1                                          & stress-induced-phosphoprotein 1                            & \num{1.8e-4}                              & 0.029                                                                           & 1.958                                      & 1.463-2.621                                & 1.957                                      & 1.451-2.640                               \\%   &                                           \\ 
\cline{1-8}
DKC1                                           & dyskeratosis congenita 1, dyskerin                         & \num{2.8e-4}                               & 0.042                                                                           & 2.046                                      & 1.490-2.808                                & 1.837                                      & 1.332-2.534                               \\%   &                                          \\ 
\cline{1-8}
%\multicolumn{1}{|l!{\color{black}\vrule}}{}    & \multicolumn{1}{l!{\color{black}\vrule}}{}                 & \multicolumn{1}{l!{\color{black}\vrule}}{} & \multicolumn{1}{l!{\color{black}\vrule}}{}                                      & \multicolumn{1}{l!{\color{black}\vrule}}{} & \multicolumn{1}{l!{\color{black}\vrule}}{} & \multicolumn{1}{l!{\color{black}\vrule}}{} & \multicolumn{1}{l!{\color{black}\vrule}}{} & \multicolumn{1}{l!{\color{black}\vrule}}{}  \\ 
%\arrayrulecolor{black}\hline

\multicolumn{8}{|l|}{\begin{tabular}[c]{@{}l@{}}Selection criteria:~~\\~Kaplan-Meier Bonferroni-adjusted \textit{P} $< 0.05 $~\\~Cox's univariate and multivariate $HR > 1.5$~\\~(* Cox's model: \textit{P} $< 0.001$)\\ \end{tabular}}   \\ 
%\cline{1-8}
%\multicolumn{8}{|l|}{Selection criteria:~\\~Kaplan-Meier Bonferroni-adjusted \textit{P} $< 0.05 $~\\~Cox's univariate and multivariate $HR > 1.5$~\\~* Cox's model: \textit{P} $< 0.001$}   \\ 
\cline{1-8}
%\multicolumn{9}{|l!{\color{black}\vrule}}{Remark: \checkmark validated by GSE2837}                i                                                                                                                                                                                                                                                                                                                                             \\
%\hline
\end{tabularx}%
} % resizebox end
\arrayrulecolor{black}
\label{table:table1}

%\end{sidewaystable}
\end{table}
%Table2. The 10 candidate genes overexpressed with better prognosis in HNSCC (ranked by Bonferroni corrected Kaplan-Meier P-value).
\begin{table}[ht]
\centering
\caption{The other top 10 genes overexpressed with better prognosis in \acrshort{hnscc} (ranked by adjusted \textit{P} value) }
\arrayrulecolor[rgb]{0.255,0.255,0.255}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|l|l|l|l|l|l|l|l|} 
\cline{1-8}
\multicolumn{1}{|c|}{\multirow{2}{*}{Gene ID}} & \multicolumn{1}{c|}{\multirow{2}{*}{Gene Description}} & \multicolumn{2}{l|}{Kaplan-Meier survival}                                                                                                                                                & \multicolumn{2}{c|}{Univariate~}                                                                        & \multicolumn{2}{c|}{Multivariate}                                                                    \\%   & \multicolumn{1}{l|}{\multirow{2}{*}{Remark}}  \\ 
\cline{3-8}
\multicolumn{1}{|c|}{}                         & \multicolumn{1}{c|}{}                                  & \multicolumn{1}{c!{\color{black}\vrule}}{\begin{tabular}[c]{@{}c@{}}FDR\\~\textit{P} value\end{tabular}}                                  & \multicolumn{1}{c!{\color{black}\vrule}}{\begin{tabular}[c]{@{}c@{}}Bonferroni\\~\textit{P} value\end{tabular}} & \multicolumn{1}{c!{\color{black}\vrule}}{HR*}   & \multicolumn{1}{c!{\color{black}\vrule}}{95\% CI}     & \multicolumn{1}{c!{\color{black}\vrule}}{HR*}   & \multicolumn{1}{c!{\color{black}\vrule}}{95\% CI}    \\%   & \multicolumn{1}{l|}{}                         \\ 
\cline{1-2}\arrayrulecolor{black}\cline{3-8}%\arrayrulecolor[rgb]{0.255,0.255,0.255}\cline{9-9}
ZNF557                                         & zinc finger protein 557                                & \multicolumn{1}{l!{\color{black}\vrule}}{\num{4.7e-6}} & \multicolumn{1}{l!{\color{black}\vrule}}{0.001}                                                      & \multicolumn{1}{l!{\color{black}\vrule}}{0.465} & \multicolumn{1}{l!{\color{black}\vrule}}{0.348-0.619} & \multicolumn{1}{l!{\color{black}\vrule}}{0.499} & \multicolumn{1}{l!{\color{black}\vrule}}{0.372-0.669}                                              \\ 
\cline{1-2}\arrayrulecolor{black}\cline{3-8}%\arrayrulecolor[rgb]{0.255,0.255,0.255}\cline{9-9}
ZNF266                                         & zinc finger protein 266                                & \num{5.2e-6}                                           & 0.001                                                                                                & 0.474                                           & 0.355-0.632                                           & 0.453                                           & 0.338-0.607                                                                                       \\ 
\hline
IL19                                           & interleukin 19                                         & \num{6.5e-6}               & 0.002                                                                                                & 0.472                                           & 0.351-0.635                                           & 0.459                                           & 0.340-0.619                                          \\%   &                                             \\ %\checkmark
\hline
MYO1H                                          & myosin 1H                                              & \num{1.4e-5}               & 0.003                                                                                                & 0.468                                           & 0.347-0.632                                           & 0.467                                           & 0.344-0.634                                          \\%   &                                              \\ 
\hline
FCGBP                                          & Fc fragment of IgG binding protein                     & \num{4.8e-5}               & 0.008                                                                                                & 0.484                                           & 0.359-0.653                                           & 0.496                                           & 0.366-0.674                                        \\%     &                                             \\ %\checkmark
\hline
%LOC148709                                      & LncRNA LOC148709                                       & \textcolor[rgb]{0,0,0.471}{\num{1.5e-6}}                                           & 0.010                                                                                                & 0.499                                           & 0.374-0.666                                           & 0.485                                           & 0.361-0.652                                           &                                              \\ 
%\hline
EVPLL                                          & envoplakin-like protein                                & \num{7.5e-5}                                           & 0.013                                                                                                & 0.490                                           & 0.363-0.661                                           & 0.494                                           & 0.364-0.672                                            \\% &                                              \\ 
\hline
PNMA5                                          & paraneoplastic antigen like 5                          & \num{3.0e-4}                                           & 0.017                                                                                                & 0.499                                           & 0.371-0.671                                           & 0.481                                           & 0.357-0.650                                            \\% &                                              \\ 
\hline
IQCN            & IQ motif containing N               & \num{1.5e-4}                                           & 0.020                                                                                                & 0.500                                           & 0.371-0.673                                           & 0.483                                           & 0.356-0.654                  \\%                           &                                              \\ %\checkmark
\hline
NPB                                            & neuropeptide B                                         & \num{2.6e-4}                                           & 0.027                                                                                                & 0.460                                           & 0.328-0.646                                           & 0.457                                           & 0.324-0.646                                            \\% &                                              \\ %\checkmark
\hline
CALML5 &
\acrlong{CALML5} &
   \num{2.0e-4} & %\num{6.54e-6} & %3.7e-7} &
   0.039 &
   0.510 &
   0.379-0.686 &
   0.493 &
   0.364-0.667  \\%  &  \\  % \checkmark \multicolumn{2}{l|} {
\hline
\multicolumn{8}{|l|}{\begin{tabular}[c]{@{}l@{}}Selection criteria:~\\~Kaplan-Meier Bonferroni-adjusted \textit{P} value \textless{} 0.05~\\~Cox's univariate and multivariate HR \textless{} 0.6\\
%\hline
%\multicolumn{9}{|l!{\color{black}\vrule}}{
(* Cox's model: \textit{P} $< 0.001$) \\
%}                                                                              
%lncRNA: Long non-coding RNA\\%Remark: \checkmark Validated by GSE2837
\end{tabular}}                                                                   \\ \hline
\end{tabular}
}
\arrayrulecolor{black}
\label{table:table3}
\end{table}

\clearpage

%%%%
